---
title: "Regression Models Course Project"
author: "Sakib Shahriar"
date: "March 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Does A Manual Car Yield A Better Miles Per Gallon Compared To An Automatic Car?


##Executive Summary
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

"Is an automatic or manual transmission better for MPG"
"Quantify the MPG difference between automatic and manual transmissions"

##Data Processing
```{r }
#load mtcars data set
data(mtcars)

#conver to factor variables
mtcars$cyl <- factor(mtcars$cyl)
mtcars$gear <- factor(mtcars$gear)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
mtcars$carb <- factor(mtcars$carb)
```

##Data Analysis

```{r }
#mpg vs auto/manual plot
boxplot(mpg~am, data = mtcars, xlab = "Transmission Type", ylab = "Miles per Gallon -MPG",
        main = "MPG by Transmission")


```

  
  
  

From the boxplot, MPG values generated from manual transmission is much better when compared to automatic.

Therefore, our hypothesis will be that transmission type is actually related to mpg values and specifically manual transmission yields better mpg than automatic. Null hyoothesis is that transmission type have no significant impact on mpg values.
  

```{r }

#So that is our hypothesis, we can use T  test now: 
tRes <-t.test(mtcars$mpg~mtcars$am)

print(tRes)

```
  
  
  
Therefore, assuming all variables remain constant, a p-value of 0.001374 which 
happens to be less than 5%, we can reject our null hypothesis and agree with the hypothesis
that indeed manual transmission yields a far better mpg.
  
  
##Quantification of the MPG difference
 
  
  First we perform a linear regression against all other variables to see the correlation that other variables may have with mpg. This is followed by a residual plot.
```{r }

#perform a linear regression against all other variables
regr <- lm(mpg~., mtcars)

compare =step(regr, direction = "both")
print(compare$coefficients)

#Residual plot
par(mfrow = c(2,2))
plot(compare)


```

##Analysis

From the coefficients table, we can see that both the cylinders have a significant negative correlation with the mpg, therefore they should be considered as a factor. Also, amManual, which is the transmission type have a strong positive correlation which backs our previous claim. Manual transmission is 1.8 times efficient compared to automatic. Weight of car also has a negative impact on mpg values, which is logical as we would expect heavier cars to be less efficient.  
From the coefficient table, it can be concluded that 


